---
title: "Lorcana TCG Forecasting: Full-Span Structural Volatility"
format: 
  html:
    theme: cosmo
    toc: true
    toc-depth: 2
    code-fold: true
    code-tools: true
execute:
  warning: false
  message: false
---

## Overview
This report evaluates the 30-day forecasting performance of a custom Single GRU model, a Multi-Window GRU Ensemble, and the pre-trained Chronos foundation model across the **entire card portfolio**. Volatility is defined by the **Lifetime Chaos Index**, measuring trend efficiency and price range.

```{r}
#| label: setup-and-prep
#| include: false

library(tidyverse)
library(scales)
library(knitr)
library(reactable)
library(lubridate)

# 1. Load Data
gru_15 <- read_csv("data/pytorch/gru_forecast_tidy_15.csv", col_types = cols(card_id = col_character())) 
gru_30 <- read_csv("data/pytorch/gru_forecast_tidy_30.csv", col_types = cols(card_id = col_character())) %>% mutate(model = "Single GRU")
gru_45 <- read_csv("data/pytorch/gru_forecast_tidy_45.csv", col_types = cols(card_id = col_character())) 

# Calculate Ensemble
ensemble_preds <- bind_rows(gru_15, gru_30, gru_45) %>%
  group_by(card_id, day_offset, actual_price) %>%
  summarize(
    pred_price = mean(pred_price), 
    conf_low = mean(conf_low), 
    conf_high = mean(conf_high),
    .groups = "drop"
  ) %>%
  mutate(model = "GRU Ensemble")

chronos_preds <- read_csv("data/chronos_forecast_tidy.csv", col_types = cols(card_id = col_character())) %>% mutate(model = "Chronos")

chronos_preds  = chronos_preds %>% drop_na(actual_price)
history_df <- read_csv("data/chronos_ready_prices.csv", col_types = cols(card_id = col_character()))

# Load Static Data and Parse the Release Date
static <- read_csv("data/target_cards_with_epids2.csv", col_types = cols(tcgplayer_id = col_character())) %>%
  mutate(released_at = mdy(released_at))

# 2. Join Predictions (NO FILTER)
combined_data <- bind_rows(gru_30, ensemble_preds, chronos_preds) %>%
  left_join(static, by = c("card_id" = "tcgplayer_id")) %>%
  mutate(card_label = paste0(name, ": ", version, " (", rarity, ")"))

# 3. Calculate Base Accuracy Metrics
accuracy_summary <- combined_data %>%
  group_by(card_id, name, version, rarity, set_name, released_at, card_label, model) %>%
  summarize(
    MAPE = mean(abs(actual_price - pred_price) / actual_price),
    .groups = "drop"
  )

# 4. Calculate LIFETIME STRUCTURAL Volatility
volatility_metrics <- history_df %>%
  filter(card_id %in% accuracy_summary$card_id) %>%
  group_by(card_id) %>%
  arrange(date) %>%
  filter(n() > 30) %>%
  slice(1:(n() - 30)) %>% 
  summarize(
    net_change = abs(last(price) - first(price)),
    sum_of_changes = sum(abs(diff(price)), na.rm = TRUE),
    efficiency_ratio = if_else(sum_of_changes > 0, net_change / sum_of_changes, 1),
    span_range = (max(price) - min(price)) / mean(price),
    volatility_cv = span_range * (1 - efficiency_ratio),
    .groups = "drop"
  ) %>%
  mutate(
    volatility_tier = ntile(volatility_cv, 3),
    volatility_label = case_when(
      volatility_tier == 1 ~ "1. Stable (Direct Trend)",
      volatility_tier == 2 ~ "2. Moderate (Minor Reversals)",
      volatility_tier == 3 ~ "3. Chaotic (Significant Reversals)"
    )
  )

# 5. Join Diagnostic Data & Calculate True Market Age
diagnostic_data <- accuracy_summary %>%
  left_join(volatility_metrics, by = "card_id") %>%
  mutate(days_since_release = as.numeric(max(history_df$date) - released_at)) %>%
  filter(!is.na(volatility_cv))

# 6. Daily Error Tracking
daily_error_data <- combined_data %>%
  left_join(volatility_metrics, by = "card_id") %>%
  filter(!is.na(volatility_cv)) %>%
  mutate(APE = abs(actual_price - pred_price) / actual_price)

# 7. Calculate Baselines
# 7. Calculate Baselines (Fixed Time Leakage)
baseline_calcs <- history_df %>%
  filter(card_id %in% unique(daily_error_data$card_id)) %>%
  group_by(card_id) %>%
  arrange(date) %>%
  # Anchor the baseline to the exact end of the model's training data
  # 30 days prior to the March 17th end-of-archive is Feb 15th.
  filter(date <= as.Date("2026-02-15")) %>% 
  summarize(
    persistence_price = last(price),
    ma_30d_price = mean(tail(price, 30)),
    .groups = "drop"
  )

baseline_daily_mape <- daily_error_data %>%
  filter(model == "Single GRU") %>% 
  left_join(baseline_calcs, by = "card_id") %>%
  mutate(
    ape_persistence = abs(actual_price - persistence_price) / actual_price,
    ape_ma = abs(actual_price - ma_30d_price) / actual_price
  ) %>%
  group_by(day_offset) %>%
  summarize(
    `Persistence` = mean(ape_persistence, na.rm = TRUE),
    `30-Day MA` = mean(ape_ma, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  pivot_longer(cols = c(`Persistence`, `30-Day MA`), names_to = "baseline_type", values_to = "mean_ape")
```

## 1. Master Accuracy Leaderboard
Use the search bar to find specific cards or sort by the **Win Margin** to see where the Single GRU outperformed the foundation model.

```{r}
#| label: tbl-master-accuracy
#| tbl-cap: "Interactive Full Portfolio Accuracy Comparison (30-Day MAPE)"

master_leaderboard <- accuracy_summary %>%
  select(card_label, rarity, set_name, model, MAPE) %>%
  pivot_wider(names_from = model, values_from = MAPE) %>%
  rowwise() %>%
  mutate(
    Winner = c("Single GRU", "GRU Ensemble", "Chronos")[which.min(c(`Single GRU`, `GRU Ensemble`, `Chronos`))],
    `Win Margin (vs Chronos)` = Chronos - `Single GRU`
  ) %>%
  ungroup()

reactable(
  master_leaderboard,
  searchable = TRUE,
  sortable = TRUE,
  striped = TRUE,
  highlight = TRUE,
  compact = TRUE,
  defaultPageSize = 10,
  columns = list(
    card_label = colDef(name = "Card Identity", minWidth = 250),
    rarity = colDef(name = "Rarity", maxWidth = 120),
    set_name = colDef(name = "Set", maxWidth = 150),
    `Single GRU` = colDef(format = colFormat(percent = TRUE, digits = 1)),
    `GRU Ensemble` = colDef(format = colFormat(percent = TRUE, digits = 1)),
    Chronos = colDef(format = colFormat(percent = TRUE, digits = 1)),
    Winner = colDef(maxWidth = 150, style = function(value) {
      color <- if (value == "Single GRU") "#e74c3c" else if (value == "Chronos") "#3498db" else "#9b59b6"
      list(color = color, fontWeight = "bold")
    }),
    `Win Margin (vs Chronos)` = colDef(
      name = "Win Margin",
      format = colFormat(percent = TRUE, digits = 1),
      style = function(value) {
        color <- if (value > 0) "#27ae60" else if (value < 0) "#c0392b" else "#2c3e50"
        list(color = color, fontWeight = "bold")
      }
    )
  )
)
```

## 2. Error by Card Market Age
Does a model need more historical data to accurately predict a card's future? This plot evaluates if newer, highly-speculative cards carry intrinsically higher forecast error than stable, older cards. Because cards are released in sets, points are horizontally jittered to display density.

```{r}
#| label: fig-error-by-age
#| fig-cap: "Relationship between a card's time on the market and its 30-Day average prediction error. Trendlines indicate model degradation on newer cards."
#| fig-width: 10
#| fig-height: 6

ggplot(diagnostic_data, aes(x = days_since_release, y = MAPE, color = model)) +
  # Use width to spread out the overlapping release dates, height = 0 to keep the exact MAPE value
  geom_jitter(alpha = 0.5, size = 2, width = 4, height = 0) + 
  geom_smooth(method = "loess", se = FALSE, size = 1.5) +
  scale_y_continuous(labels = label_percent(), limits = c(0, 0.4)) +
  scale_color_manual(values = c("Single GRU" = "#e74c3c", "GRU Ensemble" = "#9b59b6", "Chronos" = "#3498db")) +
  theme_minimal(base_size = 14) +
  theme(legend.position = "bottom") +
  labs(
    title = "Forecast Error vs. Official Market Age", 
    subtitle = "Tracking prediction accuracy against the true age of the card since release.",
    x = "Days Since Official Set Release", 
    y = "Average 30-Day MAPE", 
    color = "Model"
  )
```

## 3. Full-Span Volatility Archetypes
```{r}
#| label: fig-volatility-standardized
#| fig-cap: "Standardized Price Index (Day 1 = 1.0)."

exemplar_cards <- daily_error_data %>%
  group_by(volatility_label) %>%
  summarize(card_id = first(card_id), card_label = first(card_label), .groups = "drop")

exemplar_indexed <- history_df %>%
  filter(card_id %in% exemplar_cards$card_id) %>%
  left_join(exemplar_cards, by = "card_id") %>%
  group_by(card_id) %>%
  arrange(date) %>%
  mutate(price_index = price / first(price)) %>%
  ungroup()

ggplot(exemplar_indexed, aes(x = date, y = price_index, color = volatility_label, group = card_id)) +
  geom_line(size = 1.1) +
  geom_hline(yintercept = 1.0, linetype = "dashed", alpha = 0.5) +
  theme_minimal(base_size = 14) +
  theme(legend.position = "bottom") +
  scale_color_viridis_d(option = "viridis", end = 0.8) + 
  labs(title = "Standardized Lifetime Volatility Tiers", x = "Date", y = "Price Index", color = "Tier")
```

## 4. Forecast Degradation by Lifetime Volatility
The relationship between lifetime market chaos and model prediction failure across the forecast horizon.

```{r}
#| label: fig-forecast-degradation
#| fig-cap: "Daily error degradation segmented by lifetime volatility tiers."
#| fig-width: 10

daily_mape_summary <- daily_error_data %>%
  group_by(day_offset, model, volatility_label) %>%
  summarize(Daily_MAPE = mean(APE, na.rm = TRUE), .groups = "drop")

ggplot(daily_mape_summary, aes(x = day_offset, y = Daily_MAPE, color = model)) +
  geom_line(size = 1.2) +
  geom_point(size = 2, alpha = 0.8) +
  facet_wrap(~volatility_label, ncol = 3) +
  theme_minimal(base_size = 14) +
  theme(legend.position = "bottom") +
  scale_y_continuous(labels = label_percent()) +
  scale_color_manual(values = c("Single GRU" = "#e74c3c", "GRU Ensemble" = "#9b59b6", "Chronos" = "#3498db")) +
  labs(title = "Forecast Degradation by Lifetime Volatility", x = "Days into Future", y = "Daily MAPE", color = "Model")
```

## 5. Unmasking Uncertainty (Fan Chart)
```{r}
#| label: fig-fan-chart
#| fig-cap: "Forecast Uncertainty Horizon: Shaded bands reveal the true spread of error (75th and 90th percentiles)."
#| fig-width: 12
#| fig-height: 6

fan_data <- daily_error_data %>%
  group_by(model, day_offset) %>%
  summarize(
    median_ape = median(APE),
    p75_ape = quantile(APE, 0.75), 
    p90_ape = quantile(APE, 0.90), 
    .groups = "drop"
  )

ggplot(fan_data, aes(x = day_offset)) +
  geom_ribbon(aes(ymin = median_ape, ymax = p90_ape, fill = model), alpha = 0.2) +
  geom_ribbon(aes(ymin = median_ape, ymax = p75_ape, fill = model), alpha = 0.4) +
  geom_line(data = baseline_daily_mape, aes(x = day_offset, y = mean_ape, linetype = baseline_type), color = "#2c3e50", size = 1) +
  geom_line(aes(y = median_ape, color = model), size = 1.2) +
  facet_wrap(~ model, ncol = 3) +
  scale_y_continuous(labels = label_percent(), limits = c(0, 0.25)) +
  scale_fill_manual(values = c("Single GRU" = "#e74c3c", "GRU Ensemble" = "#9b59b6", "Chronos" = "#3498db")) +
  scale_color_manual(values = c("Single GRU" = "#c0392b", "GRU Ensemble" = "#8e44ad", "Chronos" = "#2980b9")) +
  scale_linetype_manual(values = c("Persistence" = "dotted", "30-Day MA" = "longdash")) +
  theme_minimal(base_size = 14) +
  theme(legend.position = "bottom", legend.box = "vertical") +
  labs(title = "Full Portfolio: Models vs. Statistical Baselines", x = "Days into Future", y = "APE")
```