---
title: "Market Physics of Enchanted Lorcana Cards"
subtitle: "A pooled softmax (multinomial logistic) inference model of 7-day price movements"
author: "Lorecaster — exploration/softmax_report"
date: last-modified
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    code-fold: true
    code-summary: "Show code"
    fig-width: 8
    fig-height: 4.8
    fig-dpi: 150
    theme: cosmo
execute:
  warning: false
  message: false
  cache: true
---

```{r setup}
#| cache: false
library(DBI);   library(duckdb)
library(dplyr); library(tidyr);  library(purrr)
library(nnet);  library(MASS,  exclude = "select")
library(broom); library(ggplot2)
library(knitr); library(patchwork)

# ---- chart tokens (reference dataviz palette, light mode) -------------------
tok <- list(
  surface  = "#fcfcfb", ink = "#0b0b0b", ink2 = "#52514e", muted = "#898781",
  grid     = "#e1e0d9", baseline = "#c3c2b7",
  up       = "#2a78d6",   # diverging cool pole
  down     = "#e34948",   # diverging warm pole
  flat     = "#898781",   # neutral midpoint (mark-weight gray)
  seq_lo   = "#cde2fb", seq_hi = "#0d366b",
  accent2  = "#eb6834"    # categorical slot 8 (orange) for method contrast
)
move_cols <- c(Down = tok$down, Stagnant = tok$flat, Up = tok$up)

theme_lorecaster <- function(base_size = 12) {
  theme_minimal(base_size = base_size) +
    theme(
      plot.background   = element_rect(fill = tok$surface, colour = NA),
      panel.background  = element_rect(fill = tok$surface, colour = NA),
      panel.grid.major  = element_line(colour = tok$grid, linewidth = 0.3),
      panel.grid.minor  = element_blank(),
      axis.text         = element_text(colour = tok$ink2),
      axis.title        = element_text(colour = tok$ink2),
      plot.title        = element_text(colour = tok$ink, face = "bold"),
      plot.subtitle     = element_text(colour = tok$ink2),
      plot.caption      = element_text(colour = tok$muted, size = 9),
      strip.text        = element_text(colour = tok$ink2, face = "bold"),
      legend.text       = element_text(colour = tok$ink2),
      legend.title      = element_text(colour = tok$ink2),
      legend.position   = "top"
    )
}
set.seed(1789)
```

## 1. Motivation

`exploration/softmax.R` lays out a plan: instead of training black-box,
card-specific forecasters, fit **one pooled multinomial logistic (softmax)
regression** across the whole market and read its coefficients as *market
physics* — how listing turnover, listing composition, and supply age move the
odds that a card's price goes **Down**, stays **Stagnant**, or goes **Up** over
the next week. The script builds a single 21-day snapshot and stops before any
model is fit.

This report executes the plan and extends it in four ways:

1. **Pooled panel, not one snapshot.** The plan's "Step C" calls for fitting
   across *all* historical windows simultaneously. We build a panel of sliding
   21-day windows stepped weekly across the full listing history.
2. **A lookahead leak is closed.** The prototype computed `% graded` over the
   window `[t-21, t]` — which overlaps the outcome period `(t-7, t]`. All
   features here use only information available at the decision date `t-7`.
3. **Honest uncertainty.** The same cards recur across windows, so ordinary
   Wald standard errors are too small. We report **card-clustered bootstrap**
   confidence intervals alongside the Wald ones.
4. **The softmax choice is tested, not assumed.** Down/Stagnant/Up is ordered,
   so a proportional-odds ordinal model is the parsimonious competitor. We fit
   both and run a likelihood-ratio test of the proportional-odds restriction.

One planned term dies on contact with the data: the volume-ratio × **rarity**
interaction is unidentifiable because every tracked card is Enchanted rarity.
We substitute a volume-ratio × **price tier** interaction (differential supply
elasticity across price levels) and keep the supply-age control via each
card's set release date.

## 2. Data

Listings are scraped daily from eBay into MotherDuck. We pull three compact
tables: per-card daily aggregates, distinct listing presence per day (for
turnover), and LLM-extracted listing metadata (graded vs. raw).

```{r pull-data}
con <- dbConnect(duckdb::duckdb())
invisible(dbExecute(con, "INSTALL motherduck; LOAD motherduck;"))
invisible(dbExecute(con, "ATTACH 'md:my_db' AS my_db;"))
invisible(dbExecute(con, "USE my_db;"))

daily_agg <- dbGetQuery(con, "
  SELECT id, date_pulled, count(*) AS n_active, median(price_val) AS median_price
  FROM lorcana_active_listings GROUP BY id, date_pulled")

presence <- dbGetQuery(con, "
  SELECT DISTINCT id, item_id, date_pulled FROM lorcana_active_listings")

graded_meta <- dbGetQuery(con, "
  SELECT DISTINCT id, item_id, is_graded
  FROM llm_listing_metadata WHERE is_valid = TRUE")

dbDisconnect(con, shutdown = TRUE)

meta <- readRDS("../../data/enchanteds/enchanted_list_05_07.rds") |>
  transmute(id, name, version, set_name, released_at = as.Date(released_at))
```

### 2.1 Panel construction

Each observation is a *(card, anchor date)* pair. For anchor `t`:

* **Feature window** `[t-21, t-7]`: listing turnover (distinct `item_id`s
  added vs. removed), % of unique listings graded, change in active listing
  count, price level, and prior 14-day price momentum — all knowable at `t-7`.
* **Outcome window** `(t-7, t]`: percent change in the median listing price,
  discretised at ±5% into **Down / Stagnant / Up** (the plan's thresholds).

Anchors step backward weekly from the latest pull date, keeping only anchors
where all three boundary dates exist.

```{r build-panel}
all_dates <- sort(unique(daily_agg$date_pulled))
anchors <- as.Date(character())
t <- max(all_dates)
while (t - 21 >= min(all_dates)) {
  if (all(c(t, t - 7, t - 21) %in% all_dates)) anchors <- c(anchors, t)
  t <- t - 7
}
anchors <- sort(anchors)

build_window <- function(t) {
  t7 <- t - 7; t21 <- t - 21
  p_start <- presence |> filter(date_pulled == t21) |> distinct(id, item_id)
  p_end   <- presence |> filter(date_pulled == t7)  |> distinct(id, item_id)
  turn <- full_join(mutate(p_start, was_start = 1L),
                    mutate(p_end,   was_end   = 1L),
                    by = c("id", "item_id")) |>
    group_by(id) |>
    summarise(added   = sum(is.na(was_start) & !is.na(was_end)),
              removed = sum(!is.na(was_start) & is.na(was_end)), .groups = "drop")
  grad <- presence |>
    filter(date_pulled >= t21, date_pulled <= t7) |>
    distinct(id, item_id) |>
    inner_join(graded_meta, by = c("id", "item_id")) |>
    group_by(id) |>
    summarise(pct_graded = 100 * mean(is_graded, na.rm = TRUE), .groups = "drop")
  daily_agg |>
    filter(date_pulled %in% c(t21, t7, t)) |>
    mutate(mark = case_when(date_pulled == t21 ~ "d21",
                            date_pulled == t7  ~ "d7",
                            TRUE               ~ "d0")) |>
    dplyr::select(-date_pulled) |>
    pivot_wider(names_from = mark, values_from = c(n_active, median_price)) |>
    inner_join(turn, by = "id") |>
    left_join(grad, by = "id") |>
    mutate(anchor_date = t)
}

MIN_DEPTH <- 5   # median of <5 listings is composition noise, not price signal

panel <- bind_rows(lapply(anchors, build_window)) |>
  inner_join(meta, by = "id") |>
  mutate(
    target_pct = 100 * (median_price_d0 - median_price_d7) / median_price_d7
  ) |>
  filter(!is.na(target_pct), !is.na(pct_graded),
         n_active_d21 >= MIN_DEPTH, n_active_d7 >= MIN_DEPTH,
         n_active_d0  >= MIN_DEPTH) |>
  mutate(
    move = factor(case_when(target_pct <= -5 ~ "Down",
                            target_pct >=  5 ~ "Up",
                            TRUE             ~ "Stagnant"),
                  levels = c("Stagnant", "Down", "Up")),
    log_vol_ratio  = log((added + 1) / (removed + 1)),
    log_active_chg = log(n_active_d7 / n_active_d21),
    log_price      = log(median_price_d7),
    log_momentum   = log(median_price_d7 / median_price_d21),
    pct_graded10   = pct_graded / 10,
    card_age_yrs   = as.numeric(anchor_date - released_at) / 365.25
  )
```

The result: **`r nrow(panel)` observations** on **`r n_distinct(panel$id)`
Enchanted cards** across **`r length(anchors)` weekly anchors**
(`r format(min(anchors))` → `r format(max(anchors))`), after requiring at
least `r MIN_DEPTH` active listings at every boundary date so the median
price is a real market quote rather than one seller's whim.

```{r class-mix}
#| fig-cap: "Weekly mix of 7-day price moves. The market spends most weeks with a mild upward tilt, but every regime is well represented — the pooled model is not extrapolating from a one-sided sample."
#| fig-height: 3.6
panel |>
  count(anchor_date, move) |>
  group_by(anchor_date) |>
  mutate(share = n / sum(n)) |>
  ggplot(aes(anchor_date, share, fill = move)) +
  geom_col(width = 5.4, colour = tok$surface, linewidth = 0.6) +
  scale_fill_manual(values = move_cols, name = NULL) +
  scale_y_continuous(labels = scales::percent, expand = expansion(mult = c(0, .02))) +
  scale_x_date(date_labels = "%b %d") +
  labs(x = NULL, y = "Share of cards",
       title = "Down / Stagnant / Up mix by anchor week") +
  theme_lorecaster()
```

## 3. The pooled softmax model

With **Stagnant** as the reference state, the model estimates, for each card
$i$ at anchor $t$:

$$
\log \frac{P(\text{move}_{it}=k)}{P(\text{move}_{it}=\text{Stagnant})}
= \alpha_k + \beta_{1k}\,\text{VolRatio}_{it} + \beta_{2k}\,\text{Price}_{it}
+ \beta_{3k}\,(\text{VolRatio}\times\text{Price})_{it} + \boldsymbol{\gamma}_k' \mathbf{z}_{it}
$$

for $k \in \{\text{Down}, \text{Up}\}$, where $\mathbf{z}$ holds % graded
(per 10 pp), the log change in active listings, log 14-day momentum, and card
age in years. Exponentiated coefficients are **relative-risk ratios**: the
multiplicative change in the odds of state $k$ *versus staying Stagnant*.

```{r fit-models}
f <- move ~ log_vol_ratio * log_price + pct_graded10 + log_active_chg +
            log_momentum + card_age_yrs

m_multi <- multinom(f, data = panel, trace = FALSE)
m_null  <- multinom(move ~ 1, data = panel, trace = FALSE)
mcfadden <- 1 - as.numeric(logLik(m_multi) / logLik(m_null))

wald <- tidy(m_multi, exponentiate = TRUE, conf.int = TRUE)
```

### 3.1 Is the softmax even justified? Testing the ordinal alternative

Down < Stagnant < Up is an ordered scale, so the natural null model is a
proportional-odds (ordinal logistic) regression — half the parameters, one
slope per predictor. The multinomial model nests it, so a likelihood-ratio
test asks whether predictors are allowed to push the *Down* and *Up* odds
asymmetrically.

```{r po-test}
panel$move_ord <- factor(panel$move, levels = c("Down", "Stagnant", "Up"),
                         ordered = TRUE)
m_ord <- polr(update(f, move_ord ~ .), data = panel, Hess = TRUE)
lr_po <- 2 * (logLik(m_multi) - logLik(m_ord))
df_po <- attr(logLik(m_multi), "df") - attr(logLik(m_ord), "df")
p_po  <- pchisq(as.numeric(lr_po), df_po, lower.tail = FALSE)

tibble(
  Model = c("Proportional odds (polr)", "Multinomial softmax (multinom)"),
  Parameters = c(attr(logLik(m_ord), "df"), attr(logLik(m_multi), "df")),
  LogLik = round(c(as.numeric(logLik(m_ord)), as.numeric(logLik(m_multi))), 1),
  AIC = round(c(AIC(m_ord), AIC(m_multi)), 1)
) |> kable(caption = sprintf(
  "Proportional-odds LR test: χ² = %.1f on %d df, p = %.2g — the restriction is firmly rejected.",
  as.numeric(lr_po), df_po, p_po))
```

The proportional-odds restriction fails decisively (p `r format.pval(p_po, digits = 2)`),
and AIC agrees despite the softmax spending `r df_po` extra parameters. The
market is **asymmetric**: the forces that push a card down are not mirror
images of the forces that push it up — momentum, in particular, loads almost
entirely on the *Down* equation (§4). The plan's choice of a softmax over a
simpler ordinal model is vindicated by the data, not just by taste.

### 3.2 Honest standard errors: card-clustered bootstrap

Each card contributes up to `r length(anchors)` rows, and a card's windows
share persistent traits, so rows are not independent. We resample **entire
cards** with replacement (500 replicates) and refit, which preserves
within-card dependence.

```{r cluster-boot}
B <- 500
cards <- unique(panel$id)
rhs <- ~ log_vol_ratio * log_price + pct_graded10 + log_active_chg +
         log_momentum + card_age_yrs

boot_coefs <- map_dfr(seq_len(B), function(b) {
  bd <- panel[unlist(lapply(sample(cards, length(cards), replace = TRUE),
                            function(cid) which(panel$id == cid))), ]
  fit <- try(multinom(f, data = bd, trace = FALSE), silent = TRUE)
  if (inherits(fit, "try-error")) return(NULL)
  co <- coef(fit)
  tibble(y.level = rep(rownames(co), ncol(co)),
         term    = rep(colnames(co), each = nrow(co)),
         estimate = as.vector(co), rep = b)
})

boot_ci <- boot_coefs |>
  group_by(y.level, term) |>
  summarise(conf.low  = exp(quantile(estimate, .025)),
            conf.high = exp(quantile(estimate, .975)), .groups = "drop") |>
  mutate(method = "Cluster bootstrap (by card)")

wald_ci <- wald |>
  dplyr::select(y.level, term, conf.low, conf.high) |>
  mutate(method = "Wald (i.i.d. assumption)")

or_plot <- wald |>
  dplyr::select(y.level, term, estimate) |>
  inner_join(bind_rows(boot_ci, wald_ci), by = c("y.level", "term")) |>
  filter(term != "(Intercept)")
```

## 4. Results: what moves an Enchanted's price

```{r forest}
#| fig-cap: "Relative-risk ratios (log scale) for entering Down or Up versus staying Stagnant. Filled blue = card-clustered bootstrap 95% CI; open gray = naive Wald 95% CI. A ratio of 1 (dashed) means no effect. Clustered intervals are wider, as expected — effects whose blue interval clears 1 survive honest uncertainty."
#| fig-height: 5.2
term_labels <- c(
  log_momentum     = "14-day momentum (log)",
  log_vol_ratio    = "Listing turnover ratio (log)",
  "log_vol_ratio:log_price" = "Turnover × price tier",
  log_price        = "Price level (log $)",
  pct_graded10     = "% graded (per 10 pp)",
  log_active_chg   = "Δ active listings (log)",
  card_age_yrs     = "Card age (years)"
)
or_plot |>
  mutate(term_lab = factor(term_labels[term], levels = rev(term_labels)),
         y.level  = factor(y.level, levels = c("Down", "Up"))) |>
  ggplot(aes(x = estimate, y = term_lab, colour = method, shape = method)) +
  geom_vline(xintercept = 1, linetype = "dashed", colour = tok$baseline) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high),
                 position = position_dodge(width = .65),
                 height = 0, linewidth = 0.7) +
  geom_point(position = position_dodge(width = .65), size = 2.4,
             fill = tok$surface, stroke = 0.9) +
  scale_colour_manual(values = c("Cluster bootstrap (by card)" = tok$up,
                                 "Wald (i.i.d. assumption)"    = tok$muted),
                      name = NULL) +
  scale_shape_manual(values = c("Cluster bootstrap (by card)" = 16,
                                "Wald (i.i.d. assumption)"    = 21), name = NULL) +
  scale_x_log10() +
  facet_wrap(~ y.level, nrow = 1,
             labeller = as_labeller(c(Down = "→ Down (vs Stagnant)",
                                      Up   = "→ Up (vs Stagnant)"))) +
  labs(x = "Relative-risk ratio (log scale)", y = NULL,
       title = "Odds-ratio estimates with naive vs. clustered uncertainty") +
  theme_lorecaster()
```

```{r or-table}
wald |>
  filter(term != "(Intercept)") |>
  left_join(boot_ci |> dplyr::select(y.level, term,
                                     boot_low = conf.low, boot_high = conf.high),
            by = c("y.level", "term")) |>
  transmute(Equation = y.level, Term = term_labels[term],
            `RRR` = round(estimate, 3),
            `Wald 95% CI` = sprintf("[%.2f, %.2f]", conf.low, conf.high),
            `Boot 95% CI` = sprintf("[%.2f, %.2f]", boot_low, boot_high),
            `Wald p` = signif(p.value, 2)) |>
  arrange(Equation) |>
  kable(caption = "Relative-risk ratios vs. Stagnant. Bootstrap CIs are clustered by card.")
```

Three effects survive the clustered intervals; each maps onto a hypothesis in
the original project plan.

**Mean reversion is the dominant force.** A card whose median price rose over
the prior two weeks is far more likely to give it back: one log-unit of prior
momentum multiplies the odds of *Down* by
`r round(wald$estimate[wald$y.level=="Down" & wald$term=="log_momentum"], 2)`
while leaving *Up* odds slightly *below* one. Concretely, a card that gained
+25% over days 21→7 carries about
`r round(exp(log(1.25) * log(wald$estimate[wald$y.level=="Down" & wald$term=="log_momentum"])) , 2)`×
the Down-odds of a flat card in the following week. Spikes in thinly traded
collectibles are largely composition and hype, and they decay.

**Supply pressure suppresses rallies.** The turnover ratio — new listings per
removed listing — cuts the odds of *Up* (RRR ≈
`r round(wald$estimate[wald$y.level=="Up" & wald$term=="log_vol_ratio"], 2)`
per log-unit at the average price tier). Fresh supply doesn't reliably crash a
card, but it reliably caps it. This is the plan's core "volume dynamics"
hypothesis, confirmed on the *Up* margin rather than the *Down* one.

**A graded-heavy book is a bearish tell.** Every 10 pp of graded share raises
Down-odds by ~`r round(100*(wald$estimate[wald$y.level=="Down" & wald$term=="pct_graded10"]-1))`%.
The plan framed % graded as "player liquidation vs. investor speculation
psychology" — the data sides with liquidation: when a card's shelf fills with
slabs, sellers are cashing out.

Two hypothesized effects do **not** survive:

* **Differential supply elasticity** (turnover × price) is positive but its
  clustered CI comfortably spans 1 — we cannot yet claim expensive Enchanteds
  absorb supply differently than cheap ones.
* **Price level** raises Down-odds under Wald, but the clustered interval
  widens to include 1; likewise **card age** (younger cards look more
  Down-prone, consistent with post-release supply decay, but not decisively).

```{r drop-one}
#| tbl-cap: "Drop-one likelihood-ratio tests (both equations jointly)."
drop_terms <- c("log_momentum", "log_vol_ratio:log_price", "pct_graded10",
                "log_active_chg", "card_age_yrs")
lrt <- map_dfr(drop_terms, function(tm) {
  m_red <- multinom(update(f, as.formula(paste(". ~ . -", tm))),
                    data = panel, trace = FALSE)
  lr <- 2 * (as.numeric(logLik(m_multi)) - as.numeric(logLik(m_red)))
  dfd <- attr(logLik(m_multi), "df") - attr(logLik(m_red), "df")
  tibble(Term = term_labels[tm] %||% tm, `LR χ²` = round(lr, 1), df = dfd,
         p = signif(pchisq(lr, dfd, lower.tail = FALSE), 2))
})
kable(lrt)
```

### 4.1 Effect displays

```{r effects-fn}
predict_grid <- function(grid, coefs = coef(m_multi)) {
  mm  <- model.matrix(rhs, grid)
  eta <- mm %*% t(coefs)
  ex  <- cbind(1, exp(eta))
  p   <- ex / rowSums(ex)
  colnames(p) <- c("Stagnant", rownames(coefs))
  bind_cols(grid, as_tibble(p)) |>
    pivot_longer(all_of(c("Stagnant", "Down", "Up")),
                 names_to = "move", values_to = "prob") |>
    mutate(move = factor(move, levels = c("Down", "Stagnant", "Up")))
}

boot_ribbon <- function(grid) {
  map_dfr(unique(boot_coefs$rep), function(b) {
    cb <- boot_coefs |> filter(rep == b)
    co <- matrix(cb$estimate, nrow = 2,
                 dimnames = list(unique(cb$y.level), unique(cb$term)))
    co <- co[c("Down", "Up"), colnames(coef(m_multi)), drop = FALSE]
    predict_grid(grid, co) |> mutate(rep = b)
  }) |>
    group_by(across(all_of(names(grid))), move) |>
    summarise(lo = quantile(prob, .025), hi = quantile(prob, .975),
              .groups = "drop")
}

at_typical <- summarise(panel,
  log_vol_ratio = median(log_vol_ratio), log_price = median(log_price),
  pct_graded10 = median(pct_graded10), log_active_chg = median(log_active_chg),
  log_momentum = median(log_momentum), card_age_yrs = median(card_age_yrs))
```

```{r momentum-effect}
#| fig-cap: "Predicted state probabilities vs. prior 14-day price momentum, other features at their medians. Ribbons are 95% card-clustered bootstrap bands. The Down curve is the mean-reversion signature: the bigger the run-up, the likelier the give-back."
grid_m <- at_typical[rep(1, 81), ] |>
  mutate(log_momentum = seq(log(0.5), log(2), length.out = 81))
eff_m  <- predict_grid(grid_m)
rib_m  <- boot_ribbon(grid_m)

lab_m <- eff_m |> group_by(move) |> slice_max(log_momentum, n = 1)

ggplot(eff_m, aes(exp(log_momentum) - 1, prob, colour = move)) +
  geom_ribbon(data = rib_m |> mutate(prob = NA_real_),
              aes(x = exp(log_momentum) - 1, ymin = lo, ymax = hi, fill = move),
              alpha = .14, colour = NA, inherit.aes = FALSE) +
  geom_line(linewidth = 0.9) +
  geom_text(data = lab_m, aes(label = move), hjust = -0.12, size = 3.4,
            fontface = "bold", show.legend = FALSE) +
  scale_colour_manual(values = move_cols, name = NULL) +
  scale_fill_manual(values = move_cols, guide = "none") +
  scale_x_continuous(labels = scales::percent,
                     expand = expansion(mult = c(.02, .14))) +
  scale_y_continuous(labels = scales::percent, limits = c(0, NA)) +
  labs(x = "Price change over prior 14 days (days 21 → 7)",
       y = "P(state next 7 days)",
       title = "Mean reversion: prior gains forecast give-backs") +
  theme_lorecaster()
```

```{r supply-effect}
#| fig-cap: "Predicted probability of an Up move vs. listing turnover, at the 25th/50th/75th percentile price tiers. Supply influx suppresses rallies at every tier; the tier curves are statistically indistinguishable (the interaction's CI spans 1)."
#| fig-height: 4.2
price_q <- quantile(panel$log_price, c(.25, .5, .75))
grid_s <- expand_grid(
  log_vol_ratio = seq(-1.5, 1.5, length.out = 61),
  log_price = price_q) |>
  bind_cols(at_typical |> dplyr::select(-log_vol_ratio, -log_price) |>
              slice(rep(1, 183)))
eff_s <- predict_grid(grid_s) |>
  filter(move == "Up") |>
  mutate(tier = factor(paste0("~$", round(exp(log_price))),
                       levels = paste0("~$", round(exp(price_q)))))

ggplot(eff_s, aes(log_vol_ratio, prob, colour = tier)) +
  geom_line(linewidth = 0.9) +
  geom_text(data = eff_s |> group_by(tier) |> slice_min(log_vol_ratio, n = 1),
            aes(label = tier), hjust = 1.1, size = 3.2, fontface = "bold",
            show.legend = FALSE) +
  scale_colour_manual(values = setNames(c("#86b6ef", "#2a78d6", "#0d366b"),
                                        levels(eff_s$tier)),
                      name = "Price tier") +
  scale_x_continuous(expand = expansion(mult = c(.12, .04))) +
  scale_y_continuous(labels = scales::percent, limits = c(0, NA)) +
  labs(x = "log( listings added / listings removed ), 14-day window",
       y = "P(Up next 7 days)",
       title = "Fresh supply caps rallies across price tiers") +
  theme_lorecaster()
```

## 5. Predictive honesty

The plan explicitly frames this as an inference model, and the numbers agree:
McFadden pseudo-R² is **`r round(mcfadden, 3)`** — real structure, weak
discrimination. A temporal holdout (train on the first
`r length(anchors) - 3` anchors, test on the last 3) makes the point:

```{r holdout}
test_anchors <- tail(anchors, 3)
tr <- panel |> filter(!anchor_date %in% test_anchors)
te <- panel |> filter(anchor_date %in% test_anchors)
m_tr  <- multinom(f, data = tr, trace = FALSE)
pred  <- predict(m_tr, newdata = te)
acc   <- mean(pred == te$move)
base  <- max(table(te$move)) / nrow(te)
```

```{r confusion}
#| fig-cap: "Temporal holdout confusion (last 3 anchor weeks). The model earns its keep on the Down class — the one the coefficients say it understands — while overall accuracy barely beats the majority baseline."
#| fig-height: 3.8
conf <- as.data.frame(table(pred = pred, obs = te$move))
ggplot(conf, aes(obs, pred, fill = Freq)) +
  geom_tile(colour = tok$surface, linewidth = 1.2) +
  geom_text(aes(label = Freq,
                colour = Freq > max(Freq) * .6), size = 4, fontface = "bold") +
  scale_fill_gradient(low = tok$seq_lo, high = tok$seq_hi, name = "Count") +
  scale_colour_manual(values = c(`TRUE` = tok$surface, `FALSE` = tok$ink),
                      guide = "none") +
  coord_fixed() +
  labs(x = "Observed", y = "Predicted",
       title = "Temporal holdout confusion",
       subtitle = sprintf("Accuracy %.1f%% vs. majority baseline %.1f%%",
                          100 * acc, 100 * base)) +
  theme_lorecaster() +
  theme(plot.title.position = "plot")
```

Per-class recall on the holdout:
Down **`r sprintf("%.0f%%", 100 * mean(pred[te$move == "Down"] == "Down"))`**,
Stagnant `r sprintf("%.0f%%", 100 * mean(pred[te$move == "Stagnant"] == "Stagnant"))`,
Up `r sprintf("%.0f%%", 100 * mean(pred[te$move == "Up"] == "Up"))`.
The asymmetry mirrors the inference: the model's confident coefficients
(momentum, graded share) live in the *Down* equation, so *Down* is the state
it can actually anticipate. As a **risk flag** — "this card's recent spike is
likely to retrace" — the model is useful today; as a rally-picker it is not.

## 6. Conclusions & next steps

1. **The pooled softmax works as market physics.** Mean reversion (RRR ≈
   `r round(wald$estimate[wald$y.level=="Down" & wald$term=="log_momentum"], 1)`
   on Down-odds per log-unit momentum), supply pressure capping rallies, and
   graded-share as a liquidation tell all survive card-clustered uncertainty.
2. **The softmax beats the ordinal restriction** (LR p
   `r format.pval(p_po, digits = 2)`): up-forces and down-forces are different
   physics, which an ordinal model is structurally unable to express.
3. **Wald p-values overstate certainty** in this panel — price level and card
   age lose significance once clustering is respected. Any daily monitoring
   built on this model (the plan's "Step D") should ship with the bootstrap
   intervals, not the Wald ones.
4. **Next:** two-way clustering (card × week) once more anchors accumulate;
   a graded-only price index to separate slab and raw markets; and replacing
   the ±5% fixed thresholds with card-specific volatility-scaled bands, which
   would stop high-variance cards from monopolising the Up/Down classes.

::: {.callout-note collapse="true"}
## Reproducibility

Panel: `r nrow(panel)` rows, `r n_distinct(panel$id)` cards,
`r length(anchors)` anchors (`r format(min(anchors))` → `r format(max(anchors))`),
depth filter ≥ `r MIN_DEPTH` active listings. Bootstrap: `r B` card-cluster
replicates, seed 1789. Data: MotherDuck `my_db` (`lorcana_active_listings`,
`llm_listing_metadata`), card metadata from
`data/enchanteds/enchanted_list_05_07.rds`. Render:
`quarto render exploration/softmax_report/softmax_market_inference.qmd`.
:::
